### 2025-05-12 paper/IEEE-conference-template-062824/main.tex
paper

\documentclass[conference]{IEEEtran}
\IEEEoverridecommandlockouts
% The preceding line is only needed to identify funding in the first footnote. If that is unneeded, please comment it out.
%Template version as of 6/27/2024

\usepackage{cite}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{xcolor}
\usepackage{hyperref}

\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
\begin{document}

\title{Quantum-Guided Autoencoding for Enhanced Neutral Atom Reservoir Computing in Medical Image Classification\\
{\footnotesize \textsuperscript{*}Note: Sub-titles are not captured for https://ieeexplore.ieee.org  and
should not be used}
\thanks{Identify applicable funding agency here. If none, delete this.}
}

\author{
\IEEEauthorblockN{Nuno Batista}
\IEEEauthorblockA{\textit{Department of Informatics Engineering} \\
\textit{Faculty of Sciences and Technology, University of Coimbra}\\
Coimbra, Portugal \\
\href{mailto:nunomarquesbatista@gmail.com}{nunomarquesbatista@gmail.com}}

\and
\IEEEauthorblockN{2\textsuperscript{nd} Given Name Surname}
\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
\textit{name of organization (of Aff.)}\\
City, Country \\
email address or ORCID}
}

\maketitle

\begin{abstract}

ABSTRACT PLACEHOLDER

%============================================
% ABSTRACT
%============================================
\end{abstract}

\begin{IEEEkeywords}

Reservoir Computing, Quantum-Guided Autoencoding,
Neutral Atoms, Autoencoder, Dimensionality Reduction, 
Quantum Machine Learning, Hybrid Quantum-Classical Algorithms, 
Medical Image Classification

\end{IEEEkeywords}

%============================================
% INTRODUCTION
%============================================
\section{Introduction}

INTRODUCTION PLACEHOLDER

\subsection{Subsection Example}

SUBSECTION PLACEHOLDER

%============================================
% OVERVIEW OF QUANTUM RESERVOIR COMPUTING
%============================================
\section{Overview of Quantum Reservoir Computing}

OVERVIEW PLACEHOLDER

\subsection{Neutral Atoms for Reservoir Computing}

\subsection{Feature Encoding in an Atomic Register}

%============================================
% QUANTUM-GUIDED AUTOENCODING
%============================================
\section{Overview of Quantum-Guided Autoencoding}

\subsection{Autoencoders for Dimensionality Reduction}

\subsection{Quantum-Guided Autoencoding}

%============================================
% APPLICATION OF QUANTUM-GUIDED AUTOENCODING TO RESERVOIR COMPUTING
%============================================
\section{Application of Quantum-Guided Autoencoding to Reservoir Computing}

%============================================
% REFERENCES
%============================================
\section*{References}

\end{document}